# Data Exploration

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'survival', 'survminer', 'ggplot2', 'caret'
), 'packages.bib')
```

```{r, echo=FALSE,warning=FALSE,message=FALSE,include=FALSE}
library(survival)
library(survminer)
library(tidyverse)
library(caret)
```

## Loading Data
```{r}
data(rotterdam)
```

The data that we are going to use is called `rotterdam`, and it is a dataset that's pre-recorded in the survival package. According to the documentation of the package, the data are retrieved from the Rotterdam tumor bank, which include various anonymous information about patients with breast cancer. Below is a table of the variables in the dataset:

Variable name         |      Description
----------------------|---------------------------------------------------------
`pid`                 |      patient identifier
`year`                |      year of cancer incidence
`age`                 |      age
`meno`                |      menopausal status (0= premenopausal, 1= postmenopausal)
`size`                |      tumor size, a factor with levels `<=20`, `20-50`, `>50`
`grade`               |      tumor grade
`nodes`               |      number of positive lymph nodes
`pgr`                 |      progesterone receptors (fmol/l)
`er`                  |      estrogen receptors (fmol/l)
`hormon`              |      hormonal treatment (0=no, 1=yes)
`chemo`               |      chemotherapy
`rtime`               |      days to recurrence or last follow-up
`recur`               |      0= no recurrence, 1= recurrence
`dtime`               |      days to death or last follow-up
`death`               |      0= alive, 1= dead

From the description above, we see that there are `size` which stands for the size of the tumor, `nodes` which stands for how many lymph nodes are test cancer positive, so we have 2 criterions out of the three suggested in the background info.

## Data Wrangling

```{r}
rotterdam %>% 
  group_by(size) %>%
  summarise(number = n(), .groups = 'drop')
```

The number of lymph nodes tested positive is a important measure in the TNM system. However, right now the `nodes` variable in our dataset is numeric and makes it hard for us to make visualization. For visualization purpose, we will make a new categorical variable called `Nodes_level`. For lymph nodes tested positive, the usual medical way of classifying the severity would be:`N0` for no positive nodes; `N1` for 1-3 positive nodes; `N2` for  4-9 positive nodes; and `N3` for more than 10 nodes. We will follow this classification method.

```{r}
rotterdam <- rotterdam %>% 
  mutate(Nodes_level = ifelse(nodes == 0, "N0", 
                       ifelse(nodes >= 1 & nodes <= 3, "N1", 
                       ifelse(nodes >= 4 & nodes <= 9, "N2", 
                       ifelse(nodes >= 10, "N3", NaN)))))
rotterdam %>% 
  group_by(Nodes_level) %>%
  summarise(number = n(), .groups = 'drop')
```

Since the `grade` variable in our dataset is a numeric variable whereas we actually want to treat it as a factor, we do the following:

```{r}
rotterdam <- rotterdam %>%
  mutate(grade = as.factor(grade))
```

As we were examining through the data, we found that upon the `chemo` variable and the `hormon` variable, there are instances where patients gets both therapy or neither. So in order to explore the relationship between treatment and survival, we introduce a new variable called `Treatment`, using the `chemo` and `hormon` variables.

```{r}
rotterdam <- rotterdam %>%
  mutate(Treatment = ifelse(chemo == 1 & hormon == 0, "Chemo",
                     ifelse(chemo == 0 & hormon == 1, "Hormon", 
                     ifelse(chemo == 1 & hormon == 1, "Both", "NaN/Other Treatment")))) %>% 
  mutate(Treatment = as.factor(Treatment))
```

Note that in this manner as we try to 'merge' two binary variables into one variable with four levels, we are assuming interaction between `chemo` and `hormon`.

```{r}
rotterdam_recur <- rotterdam %>%
  filter(recur == 1) %>%
  mutate(drecurtime = dtime-rtime)
nrow(rotterdam_recur)
```

We also thought it would be of interest to invesigate how recurrnce of tumor might affect the survival of the patient. We made a new dataset called `rotterdam_recur`, which only include patients with `recur` = 1. Now the dataset contains 1518 data points, a little over the original `rotterdam` dataset. We will label the time from recurrence to death as `drecurtime` in the new data frame.

## Data visualizations and exploration

### Diagnostics vs. Survival Times

It is commonly considered that the earlier the breast cancer is detected and the earlier it is treated, the longer survival a patient might enjoy. Thus we think it is important to first look at the diagnostics before treatment and visualize their relationship with survival times.

`size` vs. `dtime`
```{r}
ggplot(data = rotterdam, aes(x = size, y = dtime, fill = size)) +
  geom_violin(alpha = 0.5)
```

`size` vs. `rtime`
```{r}
ggplot(data = rotterdam, aes(x = size, y = rtime, fill = size)) +
  geom_violin(alpha = 0.5)
```

`size` vs. `drecurtime`
```{r}
ggplot(data = rotterdam_recur, aes(x = size, y = drecurtime, fill = size)) +
  geom_violin(alpha = 0.5)
```

As we can see from the three plots above, tumor size could be an important factor that affects patients' survival time and recur time. For `size` smaller than 20, most of the patients are able to survive or encounter recurrence after roughly 3000 days. But for `size` 20-50 and >50, it's highly likely for cancer cells to recur in 500 days. However, after cancer cells have recurred, most patients could not survive over 2 years. 

`Nodes_level` vs. `dtime`
```{r}
ggplot(data = rotterdam, aes(x = Nodes_level, y = dtime, fill = Nodes_level)) +
  geom_violin(alpha = 0.5)
```

`Nodes_level` vs. `rtime`
```{r}
ggplot(data = rotterdam, aes(x = Nodes_level, y = rtime, fill = Nodes_level)) +
  geom_violin(alpha = 0.5)
```

`Nodes_level` vs. `drecurtime`
```{r}
ggplot(data = rotterdam_recur, aes(x = Nodes_level, y = drecurtime, fill = Nodes_level)) +
  geom_violin(alpha = 0.5)
```

Similarly, `nodes` is also a factor impacting the life of breast cancer patients. In fact, for patients with high `Nodes_level`, it is typically considered they are either having metastasis of the cancer or already experiencing a regional recurrence of the cancer. Thus, we could see that most patients with N2 or N3 `Nodes_level` experience recurrence shortly after treatment. However, after tumor has recurred, most patients could not survive over 2 years.

### Treatment vs. Survival Times

Next we are also going to look at the effect of different types of treatments on the survival times of breast cancer patients.

`Treatment` vs. `dtime`
```{r}
ggplot(data = rotterdam, aes(x = Treatment, y = dtime, fill = Treatment)) +
  geom_violin(alpha = 0.5)
```

`Treatment` vs. `rtime`
```{r}
ggplot(data = rotterdam, aes(x = Treatment, y = rtime, fill = Treatment)) +
  geom_violin(alpha = 0.5)
```

`Treatment` vs. `drecurtime`
```{r}
ggplot(data = rotterdam_recur, aes(x = Treatment, y = drecurtime, fill = Treatment)) +
  geom_violin(alpha = 0.5)
```

By examing the three plots above, it seems that `Treatment` would not affect patients survival time or recurrence that much. We can find that `chemo + hormon` is likely to be the one with best curative effect, that patients receiving both chemo and hormon therapy tend to have longer survival time and longer time to recurrence. And the effect of hormon therapy itself seems not that satisfying. However, after tumor have recurred, most patients do not live up to 2 years.

We also found a bi-model shape in `NaN/Other Traetment` group in `rtime` vs. `Treatment`. This may because the two peaks correponds to no treatment and other treatment separately, but currently we don't have more information investigating the true reason.

### `age` + `Treatment` vs. Survival Times

`age` + `Treatment` vs. `dtime`

As we were visualizing for `Treatment` vs. `dtime`, we found that Hormontherapy generally has a weaker effect than Chemotherapy, but we think there might be some confounding variables that leads to such conclusion. One that we discovered is `age`:

```{r}
ggplot(data = rotterdam, aes(x = age, y = dtime, color = Treatment)) +
  geom_point(alpha = 0.5)
```

`age` + `Treatment` vs. `rtime`

```{r}
ggplot(data = rotterdam, aes(x = age, y = rtime, color = Treatment)) +
  geom_point(alpha = 0.5)
```

It might be different to see from the plots right now, so we decided to make a partial plot of the full plot by filtering the patients who did not take either treatment out.

```{r}
rotterdam_new <-rotterdam %>%
  filter(Treatment != "NaN/Other Treatment") %>%
  filter(Treatment != "Both")
```

`age` + `Treatment` vs. `dtime`

```{r}
ggplot(data = rotterdam_new, aes(x = age, y = dtime, color = Treatment)) +
  geom_point(alpha = 0.5)
```

`age` + `Treatment` vs. `rtime`

```{r}
ggplot(data = rotterdam_new, aes(x = age, y = rtime, color = Treatment)) +
  geom_point(alpha = 0.5)
```

From the two plots above we could see that the patients who took either chemo therapy or hormon therapy are clearly clusterd. For the group who only took chemotherapy, most patients' `age` are located below 50 years old. For the group who only took hormontherapy, most patients' age are located above 50 years old. This is because chemotherapy might have more negative effects for patients at larger age than hormontherapy and thus would effect survival if the wrong therapy is given. Generally, hormontherapy is more friendly to elder people but chemotherapy has better effect.

### General X-year Survival Rate

A very important criterion in analysis about cancer is the 5-year survival rate. In order to examine that, we introduce a new variable called `5_year_survival`, which indicates 1 if a patients survival time is larger than 5 years and 0 vice versa.

```{r}
rotterdam <- rotterdam %>%
  mutate(dtime_Years = floor(dtime/365)) %>%
  mutate(`5_year_survival` = ifelse(dtime_Years >= 5, 1, 0))
```

Now we want to calculate the 5-year survival rate for the population in the dataset.

```{r}
rotterdam %>% 
  group_by(`5_year_survival`) %>%
  summarise(number = n(), .groups = 'drop')

2084/(898+2084)
```

And also the important 10-year survival rate.

```{r}
rotterdam <- rotterdam %>%
  mutate(dtime_Years = floor(dtime/365)) %>%
  mutate(`10_year_survival` = ifelse(dtime_Years >= 10, 1, 0))
```

Now we calculate the 10-year survival rate for the population in the dataset.

```{r}
rotterdam %>% 
  group_by(`10_year_survival`) %>%
  summarise(number = n(), .groups = 'drop')

685/(685 + 2297)
```

We can find that the 5-year survival rate for breast cancer is just fine, and around 70% of patients are able to live more than 5 years. However, the 10-year survival rate is still disappointing given the current medical level, and only around 20% patients could live more than 10 years after diagnosis. 

```{r}
rotterdam %>% 
  group_by(grade) %>%
  summarise(number = n(), .groups = 'drop')
```

However, we have to notice that in our dataset, most patients are diagnosed with stage III breast cancer and a small portion are diagnosed with stage II breast cancer, which makes their cancer pretty severe already. This would give a pessimistic calculation of 5-year/ 10-year survival rates of breast cancer patients as a whole. In fact, according to webMD.com, the overall 5-year relative survival rate for breast cancer is 90% and the 10-year breast cancer relative survival rate is 84%.

Thus the important point is that female with high risk of breast caner(family inheritance, bad life habits, etc.) should have regular physical examination, with proper screening for breast cancer. Even if diagnosed, do not panic and take treatment as soon as possible. In this way, a breast cancer patients might be able to enjoy longer survival.

Another important yet sad point is that after cancer has recurred, it does not matter what treatment a patient takes and most people do not live up to 2 years if recurred. Thus patients should be extremely careful not getting cancer recurred.