# Prediction

```{r, echo=FALSE,warning=FALSE,message=FALSE,include=FALSE}
library(survival)
library(survminer)
library(tidyverse)
library(caret)
library(hrbrthemes)
library(viridis)
library(ranger)
library(randomForestSRC)
library(LTRCtrees)
library(rpart.plot)
```

Another point of interest might be how could we predict a patients survival based on some of his diagnostics and treatment record? In order to do this, we will have to fit models(parametric/ nonparametric) and generate metric of some sort based on predictions using the model.

```{r, include = FALSE}
data(rotterdam)
```

```{r, echo=FALSE,warning=FALSE,message=FALSE,include=FALSE}
rotterdam <- rotterdam %>%
  mutate(Treatment = ifelse(chemo == 1 & hormon == 0, "Chemo", 
                     ifelse(chemo == 0 & hormon == 1, "Hormon", 
                     ifelse(chemo == 1 & hormon == 1, "Both", "NaN/Other Treatment")))) %>% 
  mutate(Treatment = as.factor(Treatment))

rotterdam <- rotterdam %>% 
  mutate(Nodes_level = ifelse(nodes == 0, "N0", 
                       ifelse(nodes >= 1 & nodes <= 3, "N1", 
                       ifelse(nodes >= 4 & nodes <= 9, "N2", 
                       ifelse(nodes >= 10, "N3", NaN))))) %>% 
  mutate(Treatment = as.factor(Treatment))

rotterdam <- rotterdam %>%
  mutate(grade = as.factor(grade)) %>%
  mutate(size = as.factor(size)) %>%
  mutate(Treatment = as.factor(Treatment))

rotterdam_recur <- rotterdam %>%
  filter(recur == 1) %>%
  mutate(drecurtime = dtime-rtime)
```

## Non-parametric methods

### LRTtrees

```{r}
rotterdam <- rotterdam %>%
  mutate(start = 0)

n <- nrow(rotterdam)
rotterdam_prediction <- rotterdam[sample(n),] # shuffle row-wise
rotterdam.train = rotterdam[floor(1:(n * 0.8)),]
rotterdam.test = rotterdam[ceiling((n * 0.8)):n,]
rotterdam.test.new <- rotterdam.test %>%
  select(nodes, Treatment, dtime, death, size, age)
```

```{r}
LTRCART.obj <- LTRCART(Surv(start, dtime, death) ~ age, rotterdam.train)
LTRCIT.obj <- LTRCIT(Surv(start, dtime, death) ~ age, rotterdam.train)
```

```{r}
prp(LTRCART.obj, type = 0, roundint=FALSE)

## Plot the fitted LTRCIT tree
plot(LTRCIT.obj)
```

### Random Survival Forest

```{r}
# cross validation, stratified on status variable to ensure that  each group (here censored, dead)
# is equally distributed over the cross-validation folds
folds <- 2 # for <nfold> cross-validation
cvIndex <- createFolds(factor(rotterdam$death), folds, returnTrain = T)
```

```{r}
# create some containers to store results
# (not reasonable for big models, for big models you may want so store intermediate results on disk)
container_model <- vector("list",length(cvIndex))
container_pred <- container_model
# run RSF with default params --------------------------------------------------
# iterate through cv-folds
for(i in 1:length(cvIndex)) {
  
  # define training / test data
  train_data <- rotterdam[cvIndex[[i]],]
  eval_data <- rotterdam[-cvIndex[[i]],]
  
  # train
  rsf <- ranger(Surv(time = dtime, event = death) ~ age + Treatment + size + nodes, data = train_data)
  
  # predict (on hold out test set)
  pred <- predict(rsf, eval_data)
  
  # store results
  container_model[[i]] <- rsf
  container_pred[[i]] <- pred
}
```

show some model summaries
```{r}
container_model
```

## Parametric methods

```{r}
Lognormal <- survreg(Surv(dtime, death) ~ age + nodes + size + Treatment, dist = "lognormal", data = rotterdam.train)
predicted <- predict(Lognormal, newdata = rotterdam.test.new)
predicted
```

```{r}
MAE(rotterdam.test$dtime, predicted)
```







```{r}
# Sample the data and create a training subset.
train <- sample(1:nrow(rotterdam), round(nrow(rotterdam) * 0.80))
# Train the model.
rotterdam.grow <- rfsrc(Surv(dtime, death) ~ Treatment + size + nodes + age, rotterdam[train, ], ntree = 100)
# Test the model.
rotterdam.pred <- predict(rotterdam.grow, rotterdam[-train , ])
# Compare the results.
print(rotterdam.grow)
print(rotterdam.pred)
```

```{r}
levels(rotterdam$Treatment)
levels(rotterdam$size)
```

```{r}
treatment <- factor(levels(rotterdam$Treatment))
size <- factor(levels(rotterdam$size))

test_case <- data.frame(Treatment = treatment[2], size = size[2], nodes=5, age=56)
predict(rotterdam.grow, test_case)
```



